\section{Attributes and Attribute-sets}

An {\tt ADCL\_Attribute} is an abstraction for a particular characteristic of
a function/implementation. Each attribute is represented by a set of possible
values.  An {\tt ADCL\_Attrset} is a group of ADCL attributes.  In the
following, we present the constructors and destructors for both ADCL objects.

\begin{verbatim}
int ADCL_Attribute_create ( int maxnvalues, int *array_of_values, 
         char ** array_of_value_names, char *attr_name, 
         ADCL_Attribute *attr);

subroutine ADCL_Attribute_create ( maxnvalues, array_of_values, 
                attr, ierror )
integer maxnvalues, attr, ierr
integer array_of_values (*)
\end{verbatim}
with
\begin{itemize}
\item {\tt maxnvalues}(IN): number of possible values for this attribute
\item {\tt array\_of\_values}(IN): integer array of size {\tt maxnvalues} containing the possible values for this 
  attribute. The values in this array have to be monotony increasing. However, they do not have to be contiguous.
\item {\tt array\_of\_value\_names}(IN): character array specifying a name for
  each attribute value. Not available in the fortran interface as of now.
\item {\tt attr\_name}(IN): name of the attribute. Not available in the
  fortran interface as of now.
\item {\tt attr}(OUT): handle to the attribute object
\end{itemize} 	

\hspace{1cm}
\begin{verbatim}
int ADCL_Attribute_free   ( ADCL_Attribute *attr );

subroutine ADCL_Attribute_free ( attr, ierror )
integer attr, ierror
\end{verbatim}
with
\begin{itemize}
\item {\tt attr}(INOUT): handle to the ADCL attribute to be freed. The handle has to be a valid attribute.
    	Upon successful completion, the handle will be set to {\tt ADCL\_ATTRIBUTE\_NULL}.
\end{itemize}

\begin{verbatim}
int ADCL_Attrset_create ( int maxnum, 
         ADCL_Attribute *array_of_attrs, ADCL_Attrset *attrset );

subroutine ADCL_Attrset_create ( maxnum, array_of_attrs, 
                attrset, ierror )
integer maxnum, attrset, ierror
integer array_of_attrs (*)
\end{verbatim}
with
\begin{itemize}
\item {\tt maxnum}(IN): number of attributes to be grouped
\item {\tt array\_of\_attrs}(IN): array of size {\tt maxnum} containing the handles to attributes. Each entry of the array has to be a valid ADCL attribute created with {\tt ADCL\_Attribute\_create}.
\item {\tt attrset}(OUT): handle to the attribute-set object
\end{itemize}
\hspace{1cm}
\begin{verbatim}
int ADCL_Attrset_free ( ADCL_Attrset *attrset );

subroutine ADCL_Attrset_free ( attrset, ierror )
integer attrset, ierror
\end{verbatim}
with
\begin{itemize}
\item {\tt attrset}(INOUT): handle to the ADCL attribute-set to be freed. The handle has to be a valid attribute-set.
    	Upon successful completion, the handle will be set to {\tt ADCL\_ATTRSET\_NULL}.
\end{itemize}